---
title: "LDA Final Data Analysis"
author: "Shan Jiang"
date: "11/20/2019"
output:
    pdf_document: 
      highlight: kate
      number_sections: yes
      toc: yes
      toc_depth: 5
---

## Tool: Extract R code from Rmd document

```{r message=FALSE, warning=FALSE, paged.print=FALSE}
library(ggplot2)
library(psych) 
library(gmodels)
library(tidyr)
library(lattice)
library(timeSeries)
library(geepack)
library(knitr)
```

#### For extraction and Latex compile

```{r}
# purl("./P8157_final_sj2921.Rmd")
```

## Import data

### H1

Conduct data clean before Exploratory Data Analysis (EDA). 

```{r message=FALSE, warning=FALSE}
## Set path and read in data
getwd()
setwd("/Users/shan/Downloads/LDA/final Project/")
trial.df <- read.delim("./ALCDEP.txt",sep = "",  header = T)
colnames(trial.df) 
```

```{r}
head(trial.df)
dim(trial.df)

trial.df = trial.df %>% 
  janitor::clean_names() %>% 
  dplyr::mutate_at(., vars(treatment, gender), factor)

trial.df = trial.df %>% 
  dplyr::mutate(group1 = ifelse(treatment == 1, 1, 0)) %>% 
  dplyr::mutate(group2 = ifelse(treatment == 2, 1, 0)) %>% 
  dplyr::mutate(group3 = ifelse(treatment == 3, 1, 0)) %>% 
  dplyr::mutate_at(., vars(group1, group2, group3), factor)

str(trial.df)

```

### Visualization of time change 
```{r}
library(ggridges)
library(cyclical)
library(hrbrthemes)
ggplot(trial.long,
       aes(x = drink,y= time,fill = drink))+
  geom_density_ridges(
    aes(point_color = time, point_fill = time, 
        point_shape = time ),
    alpha = .01, point_alpha = 0.6, 
    jittered_points = TRUE) +
  scale_point_color_hue(l = 40) +
  scale_discrete_manual(aesthetics = "point_shape", values = c(21, 22, 23)) + labs(title='No. of Drinks by time for patients',
       subtitle='classified by gender') +
  theme(axis.title.y=element_blank(),
        axis.ticks.y=element_blank(),
        strip.text.y = element_text(angle = 180, hjust = 1))
```


### Bin plot
```{r}
ggplot(trial.long,
       aes(x = drink,y= time,point_color = drink,
           fill = drink))+
   geom_density_ridges_gradient(aes(fill = gender),
        scale = 3, stat = "binline",binwidth = 1, 
                                rel_min_height = 0.01) +
   scale_y_discrete(
    expand = c(0.01, 0), name = "Timepoints",
    labels = c("Baseline", "ND30", "ND60")
  )  +labs(fill = "Gender") +
  scale_x_continuous(expand = c(0, 0), 
                     name = "No. of drinks in three timepoints") +
  scale_fill_manual(values = c("#D55E0050", "#0072B250"), labels = c( "male", "female")) +
  scale_color_manual(values = c("#D55E00", "#0072B2"), guide = "none") +
  scale_discrete_manual("point_color", values = c("#D55E00", "#0072B2"), guide = "none") +
  labs(title='No. of Drinks by time for patients',
       subtitle='classified by gender') +
 theme_ridges(font_size = 13, grid = F) + theme_ridges(center = T)
 

```



#### 1.Segment Analysis: Profiles have equal levels  

Is there evidence to suggest that the treatments differ in their effects on alcohol dependence, as reflected by the number of drinks consumed in a given 30 day period?


The data format for `MANOVA` is slightly different than we saw in ANOVA. R needs each independent variable in its own vector of factors

##### 1. Test of Assumption 

1.1 Make Subcategory Test for Treatment

```{r}
group1res_nd0=lm(trial.df$nd0~ trial.df$group1)$residuals
group2res_nd0=lm(trial.df$nd0~ trial.df$group2)$residuals
group3res_nd0=lm(trial.df$nd0~ trial.df$group3)$residuals

## Normality Tests
shapiro.test(trial.df$nd0)
shapiro.test(group1res_nd0)
shapiro.test(group2res_nd0)
shapiro.test(group3res_nd0)
```

there is evidence that the data tested are not normally distributed.

1.2 For gender 

```{r}
group1res_gen=lm(trial.df$nd0~ trial.df$gender)$residuals

group2res_gen=lm(trial.df$nd30~ trial.df$gender)$residuals
group3res_gen=lm(trial.df$nd60~ trial.df$gender)$residuals
## Normality Tests
shapiro.test(group1res_gen)
shapiro.test(group2res_gen)
shapiro.test(group3res_gen)
```

1.3  Check boxplot by group 1 /Not 
```{r}
## Check boxplot 
boxplot(lm(trial.df$nd30~trial.df$group1)$residuals~trial.df$group1) # For CATEGORY
boxplot(lm(trial.df$nd60~trial.df$gender)$residuals~trial.df$gender) # For GENDER
```

1.4 check homoscedasticity

```{r}
bartlett.test(group1res_nd0~trial.df$treatment)
bartlett.test(group2res_nd0~trial.df$gender)
bartlett.test(group3res_nd0~trial.df$treatment)
```

MANOVA are both fairly robust to deviations from normality and homoscedasticity, especially when sample sizes are equal (which they are, in this case). Now, we can run the actual test 

#### 2.Run the Hypothesis Tests

```{r}
library(profileR)

mod <- pbg(trial.df[,4:6], trial.df[, 2], 
              profile.plot = TRUE)
print(mod)
```

The print function displays the average scores on the items for the 3 groups. It can be seen that the means of the drinking number are not quite similar between the three treatments. 

```{r}
res =summary(mod)
resulth2 = res$`Ho: Profiles have equal levels`
```

```{r}
library(sjPlot)
tab_df(resulth2, 
title = "Multivariate Test Results for Parallel Hypothesis", #always give
#your tables
#titles
file = "./h1_res.doc")
```


* The output shows that assuming $\alpha$ = .05, all three hypotheses  were rejected. 

* The first `Profiles are parallel` is also rejected, the results suggest that the profiles of three groups are not parallel in drinking level changes.


### H2 parallel

**H0**: The profiles in the three groups are parallel.

```{r}
resulth22 = res$`Ho: Profiles are parallel`

tab_df(resulth22, 
title = "Multivariate Test Results for Equal-level Hypothesis", #always give
file = "./h2_res3.doc")
```

Not coincidental.

### H3 Gender 

```{r}
## Pearson Chi-Squared statistic
chisq.test(table(trial.df$relapse,trial.df$gender))
```

```{r}
trial.long = gather(trial.df, time, drink, nd0:nd60) %>% 
  dplyr::select(-c(group1)) %>% 
  dplyr::mutate_at(., vars(treatment, gender, group2, group3, time), factor) %>% 
  dplyr::arrange(sid) ## Sort the obs before modelling

# Load the dataset
head(trial.long)
str(trial.long)
```

##### 1.1 -2.1 Model with only interaction (time by group)

```{r}
mf1 = drink ~ time + treatment +
              time * treatment 
              

model1  <- geeglm(mf1, 
                id=sid, 
                family= "poisson",
                corstr="ar1", 
                data=trial.long)

res1 = summary(model1)


tab_model(model1 , 
       digits = 2,
       title = "GEE model with time and treatment interaction effects", file = "./h1_model.doc")

broom::tidy(model1)
```


##### 3.1 Independence 

```{r}
mf = drink ~ time + treatment +
              time * treatment + gender +
             gender * treatment

m_ind <- geeglm(mf, 
                id=sid, 
                family= poisson(link = "log"),
                corstr="independence", 
                data= trial.long)
summary(m_ind)
```


##### 3.2 Autoregression 1

```{r}
m_ar1 <- geeglm(mf, 
                id=sid, 
                family= poisson(link = "log"),
                corstr="ar1", 
                data=trial.long)
summary(m_ar1)

```

##### 3.3  exchangeable

```{r}
m_exc <- geeglm(mf, 
                id=sid, 
                family= poisson(link = "log"),
                corstr="exchangeable", 
                data=trial.long)
summary(m_exc)
```

##### 3.4  unstructured

```{r}
m_uns <- geeglm(mf,
                id=sid, 
                family = poisson(link = "log"),
                corstr ="unstructured", 
                data=trial.long)
summary(m_uns)
```


```{r}
summary(m_ar1)

m_ar2 <- geeglm(drink ~ time + gender + group1 + group2, 
                id=sid, 
                family= poisson(link = "log"),
                corstr="ar1", 
                data=trial.long)
summary(m_ar2)

## tab results in the docx.
broom::tidy(m_ar1, file = "./h3_ar1_compare_model.doc")

tab_dfs(m_ar1,
        m_ind,
        m_exc, 
        file = "./h3_res_compare_model.doc")

require(lmtest)
lrtest(m_ar2, m_ar1)
anova(m_ar2, m_ar1)

```

### H4: Interaction 

Select the Autoregressive structure of order 1, AR(1).

```{r}
m_ar1 <- geeglm(mf, 
                id=sid, 
                family= "poisson",
                corstr="ar1", 
                data=trial.long)
summary(m_ar1)
res_gee = summary(m_ar1)
res_gee$coefficients
```


### H5: GLM -relapse
```{r}
ggplot(trial.long, aes(x = time, 
                       y = drink)) +
    geom_point(aes(color = treatment),
        size = 0.5) +
    stat_smooth(method = 'lm',
        formula = y~x,
        se = TRUE,
        aes(color = time)) +
    theme_classic()
```

```{r}
ggplot(trial.long, aes(relapse, fill = gender))  +geom_bar(position = "fill") +
    theme_classic()
```

```{r}
library(broom)
library(glmnet)

mf.2.1 = relapse ~  gender + treatment  gender *  treatment 

  
m_glm1 <- glm(mf.2.1, 
       family= binomial(link = "logit"), 
                data= trial.long)

summary(m_glm1)

tab_model(m_glm1, file =  "./h6_res.doc")

tidy(m_glm1)
```

```{r}
## The residual deviance of 29.92 on 10 d.f. is not significant:
1-pchisq(983, 936)

## at the conventional five per cent level, so we have no evidence against this model.
```

```{r}

```




